---
title: "13-anova"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#install.packages("multcomp")
library(magrittr)
library(dplyr)
library(ggplot2)
library(multcomp)
```

```{r}
View(cholesterol)
str(cholesterol)
tail(cholesterol)
```

```{r}
levels(cholesterol$trt) #multiple groups
```

```{r}
df <- cholesterol %>%
  group_by(trt) %>%
  dplyr::summarise(count_treatment = n(), 
   mean_rsp = mean(response, na.rm = TRUE),
   sd_rsp = sd(response, na.rm = TRUE)
)
```

```{r}
ggplot(cholesterol, aes(x = trt, y = response, fill = trt)) +
  geom_boxplot() +
  geom_jitter(shape = 15,
              color = c("#1b9e77"),
  position = position_jitter()) +
  theme_classic()
```

#syntax
#y ~ X1+ X2+...+Xn # X1 +  X2 +... refers to the independent variables
#y ~ . # use all the remaining variables as independent variables
```{r}
aov_model <- aov(response ~ trt)
```

#what is the average value of response?
```{r}
summary(aov_model)  #Print the summary of the test
```

#check for normality by analyzing the model results
```{r}
library(ggpubr)
ggqqplot(residuals(aov_model))
#if all the points fall approx. along the reference line, can assume normality
```

#Is The p-value lower than the usual threshold of 0.05?

#compute Shapiro-Wilcoxon test for normality
```{r}
shapiro.test(residuals(aov_model))
```

```{r}
cholesterol %>%
   group_by(trt) %>%
     rstatix::shapiro_test(response)
#data is normally distributed, the p-value >0.05

ggqqplot(cholesterol, "response", facet.by = "trt")

plot(aov_model,1)
```
#no evident relationships between residuals and fitted values. Assume homogeneity of variances.

#Pairwise comparison
#Used to find means that are significantly different from each other. Test whether there is a strong chance that an observed numerical change in one value is causally related to an observed change in another value. Tukey test is a way to test an experimental hypothesis.

```{r}
TukeyHSD(aov_model) #to determine which group has a different mean
```

#Two way ANOVA 
#y=x1 + x2 #quantitative variable and 2 categorical variables

#anova_two_way <- aov(x1 + x2, data = df)
#summary(anova_two_way)
```{r}
#install.packages("datarium")
library(datarium)

data(jobsatisfaction)
```

```{r}
set.seed(812)
jobsatisfaction %>% 
  sample_n_by(gender, education_level, size = 1)

jobsatisfaction %>%
  group_by(gender, education_level) %>%
  get_summary_stats(score, type = "mean_sd")

df2 <- ggboxplot(jobsatisfaction, x = "gender", y = "score", color = "education_level", palette = "jco")
df2

jobsatisfaction %>%
  group_by(gender, education_level) %>%
  identify_outliers(score)

res.aov <- jobsatisfaction %>% 
  rstatix::anova_test(score ~ gender * education_level)
res.aov
```

```{r}
#install.packages("AICcmodavg")
library(AICcmodavg)
```

#model.set <- list(one.way, two.way, interaction, blocking)
#model.names <- c("one.way", "two.way", "interaction", "blocking")

#aictab(model.set, modnames = model.names)

#checking for homoscedasticity
par(mfrow=c(2,2))
plot(two.way)
par(mfrow=c(1,1))