---
title: "More Descriptive Statistics"
author: "Bevera Solutions"
date: "1/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}

#set working directory 
setwd("C:\\Users\\redbe\\OneDrive\\Documents\\Data")

#import datasets
chi<-read.csv(file.choose())

pd<-read.csv("Parkinson.csv")
View(pd)

derm <-read.csv("dermatology.csv")
View(derm)

tobacco <- read.csv("tobacco.csv")
View(tobacco)
```

#f you need more descriptive statistics, use stat.desc()
```{r}
#install.packages("pastecs")
library(pastecs)
stat.desc(derm)
```

#Mosaic plots 
#A mosaic plot allows to visualize a contingency table of two qualitative variables:
```{r}
mosaicplot(table(derm$class, derm$inflammatory_monoluclear),
           color = TRUE,
           xlab = "Class", # label for x-axis
           ylab = "Mono" # label for y-axis
)
```
# a mosaic plot can also be done via the mosaic() function from the "vcd" package.
```{r}
library(vcd)
mosaic(~ class + inflammatory_monoluclear,
       data = derm,
       direction = c("v", "h")
)
```

#Barplots can only be done on qualitative variables
#draw a barplot of the relative frequencies instead of the frequencies by adding prop.table()
```{r}

barplot(prop.table(table(derm$class)))
```
#By groups
#For some statistical tests, the normality assumption is required in all groups. One solution is to draw a QQ-plot for each group by manually splitting the dataset into different groups and then draw a QQ-plot for each subset of the data (with the methods shown above). Another (easier) solution is to draw a QQ-plot for each group automatically with the argument groups = in the function qqPlot() from the {car} package
```{r}
library(car)
qqPlot(derm$class, groups = derm$itching)

# qqplot in ggplot2
library(ggplot2)
qplot(
  sample = class, data = derm,
  col = as.factor(itching),
  xlab="Itching",
  ylab="class"
)
```
#advanced descriptive stats
#more functions and packages to perform more advanced descriptive statistics in R
```{r}
#install.packages("summarytool")
library(summarytools)

#freq() for frequencies tables
#ctable() for cross-tabulations
#descr() for descriptive statistics
#dfSummary() for dataframe summaries

#A combination of these 4 functions is usually more than enough for most descriptive analysis.

freq(derm$class)

#if you need information on missing values, add report.nas = FALSE
freq(derm$class,
     report.nas = F # remove NA information
)

freq(derm$class,
     report.nas = FALSE, # remove NA information
     totals = FALSE, # remove totals
     cumul = FALSE, # remove cumuls
     headings = FALSE # remove headings
)
```

#produces cross-tabulations (also known as contingency tables) for pairs of categorical variables
```{r}
ctable(
  x = derm$class,
  y = derm$itching
)
#To add total proportions, add the argument prop = "t" 
#To remove proportions altogether, add the argument prop = "n"
#provides counts
ctable(
  x = derm$class,
  y = derm$itching,
  prop = "n", # remove proportions
  totals = FALSE, # remove totals
  headings = FALSE # remove headings
)

#The above script is equivalent to table(derm$class, derm$itching) and xtabs(~ derm$class + derm$itching)

#To display results of the Chi-square test of independence, add the chisq = TRUE argument
ctable(
  x = derm$class,
  y = derm$itching,
  chisq = TRUE, # display results of Chi-square test of independence
  headings = FALSE # remove headings
)
#if the p-value is close to 0 we reject the null hypothesis of independence between the two variables.This indicates that the variables are dependent and that there is a significant relationship between the two variables.
```

#Create a contingency table for each level of a third categorical variable with to the combination of functions, stby() and ctable().
```{r}
stby(list(
  x = tobacco$smoker, # smoker and diseased
  y = tobacco$diseased
),
INDICES = tobacco$gender, # for each gender
FUN = ctable # ctable for cross-tabulation
)
```
#The descr() function produces descriptive (univariate) statistics with common central tendency statistics and measures of dispersion.
#The function accepts single vectors as well as data frames. If a data frame is provided, all non-numerical columns are ignored so you do not have to remove them yourself before running the function.
```{r}
descr(tobacco,
      headings = FALSE, # remove headings
      stats = "common" # most common descriptive statistics
)
```

#Use stats = c("mean", "sd"), stats = "fivenum" or stats = "common"
#If you have a large number of variables, add the transpose = TRUE argument for a better display.

#To compute these descriptive statistics by group
```{r}
stby(
  data = tobacco,
  INDICES = tobacco$smoker, # by Smoker
  FUN = descr, # descriptive statistics
  stats = "common" # most common descr. stats
)
```
#Data frame summaries
#The dfSummary() function generates a summary table with statistics, frequencies and graphs for all variables in a dataset.
```{r}
dfSummary(tobacco)
```
#The describeBy() function from the {psych} package allows to report several summary statistics (i.e., number of valid cases, mean, standard deviation, median, trimmed mean, mad: median absolute deviation (from the median), minimum, maximum, range, skewness and kurtosis) by a grouping variable.
```{r}
library(psych)
describeBy(
  derm,
  derm$class # grouping variable
)
```
#The aggregate() function allows to split the data into subsets and then to compute summary statistics for each. 
```{r}
aggregate(cbind(BMI, cigs.per.day) ~ smoker + gender,
  data = tobacco,
  mean
)
```

